---
title: "Data Science HW-6"
author: "Kasturi Bhamidipati"
date: "2022-12-02"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(ggridges)
library(patchwork)
knitr::opts_chunk$set(
	echo = TRUE,
	warning = FALSE,
	fig.width = 9, 
  fig.height = 7,
  out.width = "95%"
)
theme_set(theme_minimal() + theme(legend.position = "bottom"))
options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)
scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```

# Problem 1 
 
# Problem 2 

## Importing and tidying data 

We want to import the `homicides` dataset and make the necessary changes as noted in the problem.

```{r importing homicide data, message = FALSE}
homicides = 
  read_csv("./data/homicide-data.csv") %>%
  janitor::clean_names()%>%
   mutate(
    city_state = str_c(city,state, sep = ","),
    victim_age = as.numeric(victim_age), 
    solved = as.numeric(disposition == "Closed by arrest"))%>%
  filter(
    !city_state %in% c("Dallas,TX", "Phoenix,AZ", "Kansas City,MO", "Tulsa,AL"),
    victim_race %in% c("Black", "White")
    ) %>% 
  select(victim_race, victim_age, victim_sex, city_state, solved)

```

## Baltimore `glm` 

For the city of Baltimore, MD, we want to use the `glm` function to fit a logistic regression with resolved vs unresolved as the outcome and victim age, sex and race as predictors. 

```{r baltimore glm}
baltimore_glm = 
  homicides %>% 
  filter(city_state == "Baltimore,MD") %>% 
  glm(solved ~ victim_age + victim_sex + victim_race, 
      family = binomial(), 
      data = .) %>% 
  broom::tidy() %>% 
  janitor::clean_names()%>% 
  mutate(OR = exp(estimate),
         CI_lower = exp(estimate - 1.96*std_error),
         CI_upper = exp(estimate + 1.96*std_error))

baltimore_glm
```

### Men Vs. Women Adjusted OR and 95% CI
Now we want to obtain the estimate and confidence interval of the adjusted odds ratio for solving homicides comparing male victims to female victims keeping all other variables fixed.

```{r baltimore men vs women}
baltimore_glm %>%
  filter(term == "victim_sexMale") %>%
  select(OR, CI_lower, CI_upper) %>% 
  knitr::kable(digits = 3, col.names = c('Odds Ratio', 'Lower CI', 'Upper CI'))
```

Therefore, we see that the adjusted odds ratio and its 95% CI for solving homicides comparing male victims to female victims, keeping all else constant is OR (95% CI) = `r round(baltimore_glm %>% filter(term == "victim_sexMale") %>% pull(OR), 3)` (`r round(baltimore_glm %>% filter(term == "victim_sexMale") %>% pull(CI_lower), 3)`, `r round(baltimore_glm %>% filter(term == "victim_sexMale") %>% pull(CI_upper), 3)`).

## All cities `glm` 

Now we want to do the same for all cities in the dataset. 

```{r all cities glm}
homicides_glm = 
  homicides %>%
  nest(all_cities = -city_state) %>%
  mutate(all_glm = map(.x = all_cities, ~glm(solved ~ victim_age + victim_sex + victim_race,
         family = binomial(), data = .)),
         results = map(all_glm, broom::tidy))%>%
  select(-all_glm, -all_cities) %>% 
  unnest(cols = results) %>% 
  mutate(OR = exp(estimate),
         CI_lower = exp(estimate - 1.96 * std.error),
         CI_upper = exp(estimate + 1.96 * std.error)) %>% 
  filter(term == "victim_sexMale") %>% 
  select(city_state, OR, CI_lower, CI_upper)
  
homicides_table= 
  homicides_glm%>%
  knitr::kable(digits = 3, col.names = c('City, State','Odds Ratio', 'Lower CI', 'Upper CI'))

homicides_table


```

## Plot for Adjusted OR and 95% CI 

Now we want to create a plot that shows the estimated ORs and CIs for each city. We will also organize cities according to estimated OR. 

```{r OR plot, dpi=300}
homicides_or_plot= 
  homicides_glm %>%
  mutate( city_state = fct_reorder(city_state, OR)) %>%
  ggplot(aes(x = city_state, y =OR))+
  geom_point(alpha = 0.5)+
  geom_errorbar(aes(ymin = CI_lower, ymax = CI_upper),
                alpha = 0.5)+ 
  labs(title = "Estimated Odds Ratios and their 95% CIs of Solved Homicides for Each City (Comparing Male Vs. Female Victims)", 
       x = "City, State", 
       y = "Estimated ORs and 95% CIs")+
  theme(plot.title = element_text(size = 12, hjust = 0.5),
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))

homicides_or_plot
```

### About the Plot

- From the plot above, it is evident that New York, NY has the lowest adjusted odds ratio for solving homicides, comparing male victims to female victims, keeping all else constant. On the other hand, we see that Albuquerque, NM has the highest OR for the same. 

- Next, for most cities, we can observe that the OR <1, indicating that homicide cases with male victims have lower odds of being solved compared to homicide cases of female victims, keeping all else constant. 
- However, for Nashville, TN, Fresno, CA, Stockton, CA and Albuquerque, NM, we see that the OR >1, indicating that homicide cases with male victims have higher odds of being solved compared to homicide cases of female victims, keeping all else constant. 

- It also appears to me that Atlanta, GA and Richmond, VA have OR ~1, indicating that homicide cases with male victims have similar odds of being solved compared to homicide cases of female victims, keeping all else constant.

- I also notice that for some cities, the 95% CIs are very wide (for example - Albuquerque, NM and Stockton, CA). 

# Problem 3 

## Importing and cleaning `birthweight.csv`

```{r importing brithweight, message=FALSE}
birthweight = 
  read_csv("./data/birthweight.csv")%>% 
  janitor::clean_names()%>%
  mutate(babysex = as.factor(babysex),
         frace = as.factor(frace),
         malform = as.factor(malform),
         mrace = as.factor(mrace))

sum(is.na(birthweight))
```

Therefore, we cleaned the data as appropriate; converted numeric to factor where appropriate and checked for missing data using `sum(is.na())`. No missing data were found in the `birthweight.csv` dataset. 



